Community-based marine reserves in Mexico: do they work?

### 2018-12-03 docs/paper/Villasenor-Derbez.tex

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
Aceves-Bueno\(^{1,2}\), Stuart Fulton\(^{3}\), Álvin Suarez\(^{3}\),
Arturo Hernández-Velasco\(^{3}\), Jorge Torre\(^{3}\), Fiorenza
Micheli\(^{4}\)}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
of California, Santa Barbara, Santa Barbara, CA,
USA\newline \(^{2}\)Nicholas School of the Environment, Duke University,
Beaufort, NC, USA\newline \(^{3}\)Comunidad y Biodiversidad A.C.,
Guaymas, Sonora, Mexico\newline \(^{4}\)Hopkins Marine Station and
Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
Santa Barbara, Santa Barbara, CA, 93106}

@@

\title[Community-based marine reserves]{Effectiveness of community-based marine reserves in small-scale
fisheries}

@@

Coastal marine ecosystems provide livelihoods for small-scale fishers
and coastal communities around the world. Small-scale fisheries face
great challenges since they are difficult to monitor, enforce, and
manage. Combining territorial user rights for fisheries (TURF) with
no-take marine reserves to create TURF-reserves can improve the
performance of small-scale fisheries by buffering fisheries from
environmental variability and management errors, while ensuring that
fishers reap the benefits of conservation investments. In the last 12
years, 18 old and new community-based TURF-reserves gained legal
recognition thanks to a 2014 regulation; their effectiveness has not
been formally evaluated. We combine causal inference techniques and the
Social-Ecological Systems framework to provide a holistic evaluation of
community-based TURF-reserves in three coastal communities in Mexico. We
find that while reserves have not yet achieved their stated goal of
increasing the density of lobster and other benthic invertebrates, they
continue to receive significant support from the fishing communities. A
lack of clear ecological and socioeconomic effects likely results from a
combination of factors. First, local fisheries are already well managed,
and it is unlikely that reserves might have a detectable effect. Second,
some of the reserves are not large enough to protect mobile species,
like lobster. Third, some of these reserves might be too young for the
effects to show. Fourth, variable and extreme oceanographic conditions
have impacted harvested populations. However, these reserves have shaped
small-scale fishers' way of thinking about marine conservation, which
can provide a foundation for establishing additional, larger marine
reserves needed to effectively conserve mobile species.

@@

Words in text: 3714 \textbar{} Words in headers: 35 \textbar{} Words
outside text (captions, etc.): 192

>>>>

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
Aceves-Bueno\(^{1,2}\), Stuart Fulton\(^{3}\), Álvin Suarez\(^{3}\),
Arturo Hernández-Velasco\(^{3}\), Jorge Torre\(^{3}\), Fiorenza
Micheli\(^{4}\)}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
of California, Santa Barbara, Santa Barbara, CA, USA \newline
\(^{2}\)Nicholas School of the Environment, Duke University, Beaufort,
NC, USA \newline \(^{3}\)Comunidad y Biodiversidad A.C., Guaymas,
Sonora, Mexico \newline \(^{4}\)Hopkins Marine Station and Center for
Ocean Solutions, Stanford University, Pacific Grove, CA, USA" \newline}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
Santa Barbara, Santa Barbara, CA, 93106}

@@

\title[Community-based marine reserves]{Effectiveness of community-based TURF-reserves in small-scale fisheries}

@@

Coastal marine ecosystems provide livelihoods for small-scale fishers
and coastal communities around the world. Small-scale fisheries face
great challenges since they are difficult to monitor, enforce, and
manage. Combining territorial user rights for fisheries (TURF) with
no-take marine reserves to create TURF-reserves can improve the
performance of small-scale fisheries by buffering fisheries from
environmental variability and management errors, while ensuring that
fishers reap the benefits of conservation investments. In the last 12
years, 18 old and new community-based TURF-reserves gained legal
recognition thanks to a 2014 regulation; their effectiveness has not
been formally evaluated. We combine causal inference techniques and the
Social-Ecological Systems framework to provide a holistic evaluation of
community-based TURF-reserves in three coastal communities in Mexico. We
find that while reserves have not yet achieved their stated goal of
increasing the density of lobster and other benthic invertebrates, they
continue to receive significant support from the fishing communities. A
lack of clear ecological and socioeconomic effects likely results from a
combination of factors. First, some of these reserves might be too young
for the effects to show. Second, the reserves are not large enough to
protect mobile species, like lobster. Third, variable and extreme
oceanographic conditions have impacted harvested populations. Fourth,
local fisheries are already well managed, and it is unlikely that
reserves might have a detectable effect in landings. However, these
reserves have shaped small-scale fishers' way of thinking about marine
conservation, which can provide a foundation for establishing
additional, larger marine reserves needed to effectively conserve mobile
species.

@@

### 2018-12-03 slides/GainesLabPresentation.Rmd

zrp <- st_read(dsn = here::here("data", "spatial"), layer = "ZRP", quiet = T) %>% 
  sf::st_centroid() %>% 
  select(Name, geometry) %>% 
  mutate(Type = "FR")

rescom <- st_read(dsn = here::here("data", "spatial"), layer = "Res_Com", quiet = T) %>% 
  sf::st_centroid() %>% 
  select(Name = Nombre, geometry) %>% 
  mutate(Type = "CB")

@@

ggplot() +
  geom_sf(data = W) +
  geom_sf(data = reserves, aes(fill = Type), shape = 21) +
  theme_classic() +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = c(0.9, 0.5)) +
  lims(x = c(-118.36648, -86.71074), y = c(14.53507, 32.71863))
```

>>>>

zrp <- st_read(dsn = here::here("raw_data", "spatial"), layer = "ZRP", quiet = T) %>% 
  sf::st_centroid() %>% 
  select(Name, geometry) %>% 
  mutate(Type = "FR")

rescom <- st_read(dsn = here::here("raw_data", "spatial"), layer = "Res_Com", quiet = T) %>% 
  sf::st_centroid() %>% 
  select(Name = Nombre, geometry) %>% 
  mutate(Type = "CB")

@@

ggplot() +
  geom_sf(data = W) +
  geom_sf(data = reserves, shape = 21, fill = "red", color = "black") +
  theme_classic() +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = c(0.9, 0.5)) +
  lims(x = c(-118.36648, -86.71074), y = c(14.53507, 32.71863))
```
